### 2007-01-15 Documentation/part1.tex
JK - Minor corrections chapter 11, modified documentation of glbGetChiFunction

{\tt int glbGetChiFunction(int exp, int rule, int on\_off, char *sys\_id)}

@@

systematics on or off as given by {\tt on\_off}. The user must 
ensure that the return string {\tt sys\_id} is large enough to hold the name of the function.

>>>>

{\tt int glbGetChiFunction(int exp, int rule, int on\_off, char *sys\_id, size\_t max\_len)}

@@

systematics on or off as given by {\tt on\_off}. The name is copied to the string
{\tt sys\_id}, the maximum length of which is specified by {\tt max\_len}. If
{\tt max\_len} is too small, or if any other error occurs, the return value is $< 0$.

### 2007-01-15 Documentation/part2.tex
JK - Minor corrections chapter 11, modified documentation of glbGetChiFunction

As an argument, {\tt globes} takes a {\tt .glb}-file. While parsing it,
it prints any warnings and errors which have occured during reading the file. Then it uses the experiment description in the file to compute the event rates at a certain point in parameter space. Finally, it displays the result based on the options used to call {\tt globes}. 
The options of {\tt globes} follow the GNU standard. Thus, there
is a {\tt \verb+--+help} option to display all other options 
together with short descriptions.

\index{norm}{globes@{\tt globes}!oscillation parameters}
Calling {\tt globes} without any options and with a {\tt .glb}-file as argument produces an event summary at rule level. In this case,  
the full experiment description in the file is taken into account, \ie, all efficiencies, backgrounds, and energy resolution effects. 
Thus, the returned event rates are the ones which will be actually 
used to compute the $\chi^2$ later. By default, the oscillation parameters used to calculate the transition probability are
\begin{eqnarray}
\label{eq:globes_params}
\sin^22\theta_{12}=0.8&\quad&\Delta m^2_{21}=7\cdot10^{-5}
\,\mathrm{eV}^2\,,\nonumber\\
\sin^22\theta_{23}=1.0&\quad&\Delta m^2_{31}=3\cdot10^{-3}
\,\mathrm{eV}^2\,,\nonumber\\
\delta=0&\quad&\sin^22\theta_{13}=0.1\,.\
\end{eqnarray}
Of course, it is possible to change these default values either by using the
option {\tt -p} on a call by call basis, or by setting the environment variable
{\tt GLB\_CENTRAL\_VALUES}:
%
\index{norm}{GLB CENTRAL VALUES@{\tt GLB\_CENTRAL\_VALUES}} 
\index{norm}{Environment variables!GLB CENTRAL VALUES@{\tt GLB\_CENTRAL\_VALUES}}
%
\begin{quote}
{\tt
globes -p'0.55,0,0.785,0,0.0008,0.0025'\\
globes \verb+--+parameters='0.55,0,0.785,0,0.0008,0.0025'
}
\end{quote}
For example, {\tt GLB\_CENTRAL\_VALUES} can be defined
within the shell session or in the shell profile:
\begin{quote}
{\tt
export GLB\_CENTRAL\_VALUES='0.55,0,0.785,0,0.0008,0.0025'
}
\end{quote}
Note that in the case of additional non-standard parameters, these will just to 
be included at the end of these lists.
Furthermore it is possible to switch off oscillations with the {\tt -N} option and to switch them on again with {\tt -O} (the default). The effect of {\tt -N} is the same as to use {\tt NOSC\_} in all oscillation channels.
Thi feature is useful if one wants to normalize an expriment flux if
the number of un-oscillated events is given.

@@

In the process of defining a new experiment, the default output of {\tt globes} at rule level is the final step. However, in order to arrive 
at this level it is often necessary to review the intermediate steps in the event rate calculation. The {\tt globes} utility offers many possibilities to do this based on the rate access functions as described in \Sec~\ref{sec:event_rates}.

@@

As a first step, one may want to check if each channel produces the 
anticipated output. Channel rates are returned if the {\tt -c} option 
is used.
This option takes as an optional argument the channel number 
(starting at zero). If no argument is given all channels are displayed.
By default, the sum of the event rates in each channel is shown. Each 
column has as first line the same channel name as in the file.

It is also possible to switch off one detector effect after the other. 
First, one can switch off the post-smearing efficiencies ({\tt -f}) and the 
post-smearing backgrounds ({\tt -g}). Next, one can switch off the 
energy resolution function with ({\tt -b}) and view the rates before
smearing. If the {\tt -s} option is also used, the number of
lines in the output will be given by {\tt \$sampling\_points}.
Another effect of the {\tt -b} option is that the post-smearing efficiencies
and backgrounds are no longer taken into account. Therefore, the {\tt -g} 
and {\tt -f} options now apply to the \emph{pre}-smearing efficiencies 
and the \emph{pre}-smearing backgrounds. Thus, 
\begin{quote}
{\tt
globes -c -b -g -f FILE
}
\end{quote}
produces the raw event rate corresponding to the convolution of flux, 
probability, and cross section, which is neglecting all detector effects.

@@

The next logical step after checking the channel rates is to investigate
 the rule rates. The rule rates are returned with the option {\tt -r}.
This option takes as an optional argument the rule number 
(starting at zero). If no argument is given, all rules will be displayed.
By default, the sum of the event rates in each rule is shown, as well
as for each component within the rule. Each 
rule is preceeded by a line with the same rule name as in the file.

It is also possible for rules to switch off one detector effect after 
the other -- with the limitation that rules only make sense
after the energy resolution function has been applied to each channel.
Therefore, it is \emph{not} possible to use {\tt -b} together with {\tt -r}, or to switch off any pre-smearing efficiencies or backgrounds. 
One can, however, switch off the post-smearing efficiencies ({\tt -f}) 
and the post-smearing backgrounds ({\tt -g}) for each channel. Since
the definition of a rule also contains so-called ``coefficients'', it is
possible to switch them off with {\tt -i}. 
% This options also deactivates
% any setting of {\tt @backgroundcenter}.

@@

There are special options for certain special formats: {\tt -m} produces
Mathematica\footnote{Mathematica is a trademark of Wolfram Inc.} list
output, which can be directly visualized by {\tt MultipleListPlot}.
The option {\tt -u} uses the same principal formatting as {\tt -m}, but
it allows to specify the left, middle, and right delimiters in constructing
the list, such as
\begin{quote}
{\tt
left\\
left 1 middle 2 middle 3 right\\
middle\\
left\\
left 1 middle 2 middle 3 right\\
right\\
}
\end{quote}
This is, with {\tt left} = '\{', {\tt middle} = ',' and {\tt right} = '\}',
equivalent to the list $\{\{1,2,3\},\{1,2,3\}\}$. 
The delimiters can be set by {\tt -L}, {\tt -M} and {\tt -R} as in the 
following example:
\begin{quote}
{\tt
globes -Su -R\$'\verb+\n+' \verb+--+Middle=" " -L" " ...
}
\end{quote}
Here {\tt \$'\verb+\n+'} is the escape sequence in the shell for ANSI C-like characters, such as linefeed '\verb+\n+'. The above example
produces a a two column file such as
\begin{quote}
1.0 0.12\\
1.2 0.14\\
1.3 0.18\\
...
\end{quote}
where the first column is the central energy of the bin or the sampling point,
 and
the second column gives the event rate. Usually, the output is a concatenation of many such two columns tables, where 
each rule part or channel part has its own table. Thus one can, by using
{\tt -u} and user-defined delimiters, construct many different  
output formats.

>>>>

As an argument, {\tt globes} takes a {\tt .glb}-file. While parsing it,
it prints any warnings and errors which occur while reading the file. Then it uses the experiment description in the file to compute the event rates at a certain point in parameter space. Finally, it displays the result based on the options used to call {\tt globes}. 
The options of {\tt globes} follow the GNU standard. Thus, there
is a {\tt \verb+--+help} option to display all other options 
together with short descriptions.

\index{norm}{globes@{\tt globes}!oscillation parameters}
Calling {\tt globes} without any options and with a {\tt .glb}-file as argument produces an event summary at rule level. In this case,  
the full experiment description in the file is taken into account, \ie, all efficiencies, backgrounds, and energy resolution effects. 
Thus, the returned event rates are the ones which will be actually 
used to compute the $\chi^2$ later. By default, the oscillation parameters used to calculate the transition probability are
\begin{eqnarray}
\label{eq:globes_params}
\sin^22\theta_{12}=0.8&\quad&\Delta m^2_{21}=7\cdot10^{-5}
\,\mathrm{eV}^2\,,\nonumber\\
\sin^22\theta_{23}=1.0&\quad&\Delta m^2_{31}=3\cdot10^{-3}
\,\mathrm{eV}^2\,,\nonumber\\
\delta=0&\quad&\sin^22\theta_{13}=0.1\,.\
\end{eqnarray}
Of course, it is possible to change these default values either by using the
option {\tt -p} on a call by call basis, or by setting the environment variable
{\tt GLB\_CENTRAL\_VALUES}:
%
\index{norm}{GLB CENTRAL VALUES@{\tt GLB\_CENTRAL\_VALUES}} 
\index{norm}{Environment variables!GLB CENTRAL VALUES@{\tt GLB\_CENTRAL\_VALUES}}
%
\begin{quote}
{\tt
globes -p'0.55,0,0.785,0,0.0008,0.0025'\\
globes \verb+--+parameters='0.55,0,0.785,0,0.0008,0.0025'
}
\end{quote}
For example, {\tt GLB\_CENTRAL\_VALUES} can be defined
within the shell session or in the shell profile:
\begin{quote}
{\tt
export GLB\_CENTRAL\_VALUES='0.55,0,0.785,0,0.0008,0.0025'
}
\end{quote}
Note that in the case of additional non-standard parameters, these will have to 
be included at the end of these lists.
Furthermore it is possible to switch off oscillations with the {\tt -N} option and to switch them on again with {\tt -O} (the default). The effect of {\tt -N} is the same as to use {\tt NOSC\_} in all oscillation channels.
This feature is useful if one wants to normalize the flux in an experimental to
a given number of un-oscillated events.

@@

In the process of defining a new experiment, the default output of {\tt globes} at rule level is the final step. However, in order to arrive 
at this level it is often necessary to review the intermediate steps in the event rate calculation. The {\tt globes} utility offers many possibilities to do this based on the rate access functions described in \Sec~\ref{sec:event_rates}.

@@

As a first step, one may want to check if each channel produces the 
anticipated output. Channel rates are returned if the {\tt -c} option 
is used.
This option takes as an optional argument the channel number 
(starting at zero). If no argument is given, all channels are displayed.
By default, the sum of the event rates in each channel is shown. Each 
column has as first line the same channel name as in the file.

It is also possible to switch off one detector effect after the other. 
First, one can switch off the post-smearing efficiencies ({\tt -f}) and the 
post-smearing backgrounds ({\tt -g}). Next, one can switch off the 
energy resolution function with ({\tt -b}) and view the rates before
smearing. If the {\tt -s} option is also used, the number of
lines in the output will be given by {\tt \$sampling\_points}.
Another effect of the {\tt -b} option is that the post-smearing efficiencies
and backgrounds are no longer taken into account. Therefore, the {\tt -g} 
and {\tt -f} options now apply to the \emph{pre}-smearing efficiencies 
and the \emph{pre}-smearing backgrounds. Thus, 
\begin{quote}
{\tt
globes -c -b -g -f FILE
}
\end{quote}
produces the raw event rate corresponding to the convolution of flux, 
probability, and cross section, neglecting all detector effects.

@@

The next logical step after checking the channel rates is to investigate
 the rule rates. The rule rates are returned with the option {\tt -r}.
This option takes as an optional argument the rule number 
(starting at zero). If no argument is given, all rules will be displayed.
By default, the signal and background rates in each rule are shown
separately, together with their decomposition into the different
contributing channels. Each rule is preceeded by a line with the same
rule name as in the file.

Also for the rules, it is possible to switch off one detector effect after 
the other -- with the limitation that rules only make sense
after the energy resolution function has been applied to each channel.
Therefore, it is \emph{not} possible to use {\tt -b} together with {\tt -r}, or to switch off any pre-smearing efficiencies or backgrounds. 
One can, however, switch off the post-smearing efficiencies ({\tt -f}) 
and the post-smearing backgrounds ({\tt -g}) for each channel. Since
the definition of a rule also contains so-called ``coefficients'', it is
possible to switch them off with {\tt -i}. 
% This options also deactivates
% any setting of {\tt @backgroundcenter}.

@@

There are special options for certain special formats: {\tt -m} produces
Mathematica\footnote{Mathematica is a trademark of Wolfram Inc.} list
output, which can be directly visualized by {\tt MultipleListPlot}.
The option {\tt -u} uses in principle the same formatting as {\tt -m}, but
it allows to specify the left, middle, and right delimiters in constructing
the list, such as
\begin{quote}
{\tt
left\\
left 1 middle 2 middle 3 right\\
middle\\
left\\
left 1 middle 2 middle 3 right\\
right\\
}
\end{quote}
This is, with {\tt left} = '\{', {\tt middle} = ',' and {\tt right} = '\}',
equivalent to the list $\{\{1,2,3\},\{1,2,3\}\}$. 
The delimiters can be set by {\tt -L}, {\tt -M} and {\tt -R} as in the 
following example:
\begin{quote}
{\tt
globes -Su -R\$'\verb+\n+' \verb+--+Middle=" " -L" " ...
}
\end{quote}
Here {\tt \$'\verb+\n+'} is the escape sequence in the shell for ANSI C-like characters, such as linefeed '\verb+\n+'. The above example
produces a a two column file such as
\begin{quote}
1.0 0.12\\
1.2 0.14\\
1.3 0.18\\
...
\end{quote}
where the first column is the central energy of the bin or the sampling point,
 and
the second column gives the event rate. Usually, the output is a concatenation of many such two columns tables, where 
each rule part or channel part has its own table. Thus one can, by using
{\tt -u} and user-defined delimiters, construct many different  
output formats.
